---
title: "Top5mVersusBottom1m"
author: "Elizabeth Davenport"
date: '2023-09-14'
output: word_document
---

# Initial stuff, including loading packages and importing data

##loading packages
```{r,message=F,warning=F}
library(here)
library(rstatix)
library(ggplot2)
library(tidyverse)
library(dplyr)
```
## loading files

```{r,message=F,warning=F}
# Tell R where files are stored
here::i_am("scripts/Top5mVersusBottom1m.Rmd")

# Load Files
a <- readr::read_csv(here("data/SporeConcentrationsByDepthAll.csv"))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Walsh Ordospora 6 meters
```{r,message=F,warning=F}

wo6 = a %>%
  filter(Parasite== "Ordospora",
         Lake=="Walsh",
         Depth =="6")
wo6.spores <- wo6$Concentration

wo = a %>%
  filter(Parasite== "Ordospora",
         Lake=="Walsh",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

wo.spores<- wo$Concentration

walsh.ordo <- data.frame( 
                group = c(rep("6",7), rep("other",66)),
                spores = c(wo6.spores,  wo.spores)
                )

res.wo <- wilcox.test(wo6.spores, wo.spores)
res.wo
### significantly different, W = 371.5, p = 0.0083
```

### Walsh Metschnikowia 6 meters
```{r,message=F,warning=F}

wm6 = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="Walsh",
         Depth =="6")
wm6.spores <- wm6$Concentration

wm = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="Walsh",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

wm.spores<- wm$Concentration

walsh.metsch <- data.frame( 
                group = c(rep("6",7), rep("other",66)),
                spores = c(wm6.spores,  wm.spores)
                )

res.wm <- wilcox.test(wm6.spores, wm.spores)
res.wm
### significantly different, W = 413, p = 0.0005
```



### Walsh Blastulidium 6 meters
```{r,message=F,warning=F}

wb6 = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="Walsh",
         Depth =="6")
wb6.spores <- wb6$Concentration

wb = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="Walsh",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

wb.spores<- wb$Concentration

walsh.brood <- data.frame( 
                group = c(rep("6",7), rep("other",66)),
                spores = c(wb6.spores,  wb.spores)
                )

res.wb <- wilcox.test(wb6.spores, wb.spores)
res.wb
### not significantly different, W = 299.5, p = 0.20
```

### Walsh Pasteuria 6 meters
```{r,message=F,warning=F}

wp6 = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="Walsh",
         Depth =="6")
wp6.spores <- wp6$Concentration

wp = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="Walsh",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

wp.spores<- wp$Concentration

walsh.past <- data.frame( 
                group = c(rep("6",7), rep("other",66)),
                spores = c(wp6.spores,  wp.spores)
                )

res.wp <- wilcox.test(wp6.spores, wp.spores)
res.wp
### significantly different, W = 393.5, p = 0.0015
```

## Walsh Spirobacillus 6 meters
```{r,message=F,warning=F}

ws6 = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="Walsh",
         Depth =="6")
ws6.spores <- ws6$Concentration

ws = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="Walsh",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

ws.spores<- ws$Concentration

walsh.spiro <- data.frame( 
                group = c(rep("6",7), rep("other",66)),
                spores = c(ws6.spores,  ws.spores)
                )

res.ws <- wilcox.test(ws6.spores, ws.spores)
res.ws
### not significantly different, W = 326, p = 0.07
```

### Mill Ordospora 6 meters
```{r,message=F,warning=F}

mo6 = a %>%
  filter(Parasite== "Ordospora",
         Lake=="Mill",
         Depth =="6")
mo6.spores <- mo6$Concentration

mo = a %>%
  filter(Parasite== "Ordospora",
         Lake=="Mill",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

mo.spores<- mo$Concentration

mill.ordo <- data.frame( 
                group = c(rep("6",6), rep("other",65)),
                spores = c(mo6.spores,  mo.spores)
                )

res.mo <- wilcox.test(mo6.spores, mo.spores)
res.mo
### not significantly different, W = 218.5, p = 0.63
```

### Mill Metschnikowia 6 meters
```{r,message=F,warning=F}

mm6 = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="Mill",
         Depth =="6")
mm6.spores <- mm6$Concentration

mm = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="Mill",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

mm.spores<- mm$Concentration

mill.metsch <- data.frame( 
                group = c(rep("6",6), rep("other",65)),
                spores = c(mm6.spores,  mm.spores)
                )

res.mm <- wilcox.test(mm6.spores, mm.spores)
res.mm
### not significantly different, W = 228, p = 0.457
```

### Mill Blastulidium 6 meters
```{r,message=F,warning=F}

mb6 = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="Mill",
         Depth =="6")
mb6.spores <- mb6$Concentration

mb = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="Mill",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

mb.spores<- mb$Concentration

res.mb <- wilcox.test(mb6.spores, mb.spores)
res.mb
### significantly different, W = 300.5, p = 0.027
```

### Mill Pasteuria 6 meters
```{r,message=F,warning=F}

mp6 = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="Mill",
         Depth =="6")
mp6.spores <- mp6$Concentration

mp = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="Mill",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

mp.spores<- mp$Concentration

res.mp <- wilcox.test(mp6.spores, mp.spores)
res.mp
### significantly different, W = 273, p = 0.027
```

## Mill Spirobacillus 6 meters
```{r,message=F,warning=F}

ms6 = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="Mill",
         Depth =="6")
ms6.spores <- ms6$Concentration

ms = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="Mill",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

ms.spores<- ms$Concentration

res.ms <- wilcox.test(ms6.spores, ms.spores)
res.ms
### not significantly different, W = 326, p = 0.4
```

### Cedar Ordospora top 5 meters v. 1 meter
```{r,message=F,warning=F}

co6 = a %>%
  filter(Parasite== "Ordospora",
         Lake=="Cedar",
         Depth =="7")
co6.spores <- co6$Concentration

co = a %>%
  filter(Parasite== "Ordospora",
         Lake=="Cedar",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

co.spores<- co$Concentration

res.co <- wilcox.test(co6.spores, co.spores)
res.co
### not significantly different, W = 461, p = 0.13
```

### Cedar Metschnikowia top 5 meters v. 1 meter
```{r,message=F,warning=F}

cm6 = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="Cedar",
         Depth =="7")
cm6.spores <- cm6$Concentration

cm = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="Cedar",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

cm.spores<- cm$Concentration

res.cm <- wilcox.test(cm6.spores, cm.spores)
res.cm
### not significantly different, W = 434, p = 0.16
```

### Cedar Blastulidium top 5 meters v. 1 meter
```{r,message=F,warning=F}

cb6 = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="Cedar",
         Depth =="7")
cb6.spores <- cb6$Concentration

cb = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="Cedar",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

cb.spores<- cb$Concentration

res.cb <- wilcox.test(cb6.spores, cb.spores)
res.cb
### not significantly different, W = 415, p = 0.4
```

### Cedar Pasteuria top 5 meters v. 1 meter
```{r,message=F,warning=F}

cp6 = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="Cedar",
         Depth =="7")
cp6.spores <- cp6$Concentration

cp = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="Cedar",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

cp.spores<- cp$Concentration

res.cp <- wilcox.test(cp6.spores, cp.spores)
res.cp
### significantly different, W = 563, p = 0.0003
```

## Cedar Spirobacillus top 5 meters v. 1 meter
```{r,message=F,warning=F}

cs6 = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="Cedar",
         Depth =="7")
cs6.spores <- cs6$Concentration

cs = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="Cedar",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

cs.spores<- cs$Concentration

res.cs <- wilcox.test(cs6.spores, cs.spores)
res.cs
### not significantly different, W = 477, p = 0.07
```

### Crooked W Ordospora 6 meters
```{r,message=F,warning=F}

cwo6 = a %>%
  filter(Parasite== "Ordospora",
         Lake=="CrookedW",
         Depth =="6")
cwo6.spores <- cwo6$Concentration

cwo = a %>%
  filter(Parasite== "Ordospora",
         Lake=="CrookedW",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

cwo.spores<- cwo$Concentration

res.cwo <- wilcox.test(cwo6.spores, cwo.spores)
res.cwo
### significantly different, W = 537, p = 0.007
```

### Crooked W Metschnikowia 6 meters
```{r,message=F,warning=F}

cwm6 = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="CrookedW",
         Depth =="6")
cwm6.spores <- cwm6$Concentration

cwm = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="CrookedW",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

cwm.spores<- cwm$Concentration

res.cwm <- wilcox.test(cwm6.spores, cwm.spores)
res.cwm
### not significantly different, W = 288.5, p = 0.28
```

### Crooked W Blastulidium 6 meters
```{r,message=F,warning=F}

cwb6 = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="CrookedW",
         Depth =="6")
cwb6.spores <- cwb6$Concentration

cwb = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="CrookedW",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

cwb.spores<- cwb$Concentration

res.cwb <- wilcox.test(cwb6.spores, cwb.spores)
res.cwb
### significantly different, W = 559, p = 0.0009
```

### Crooked W Pasteuria 6 meters
```{r,message=F,warning=F}

cwp6 = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="CrookedW",
         Depth =="6")
cwp6.spores <- cwp6$Concentration

cwp = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="CrookedW",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

cwp.spores<- cwp$Concentration

res.cwp <- wilcox.test(cwp6.spores, cwp.spores)
res.cwp
### significantly different, W = 477.5, p = 0.04
```

## Crooked W Spirobacillus 6 meters
```{r,message=F,warning=F}

cws6 = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="CrookedW",
         Depth =="6")
cws6.spores <- cws6$Concentration

cws = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="CrookedW",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

cws.spores<- cws$Concentration

res.cws <- wilcox.test(cws6.spores, cws.spores)
res.cws
### not significantly different, W = 466, p = 0.11
```

### LilAp Ordospora 6 meters
```{r,message=F,warning=F}

lo6 = a %>%
  filter(Parasite== "Ordospora",
         Lake=="LilAp",
         Depth =="6")
lo6.spores <- lo6$Concentration

lo = a %>%
  filter(Parasite== "Ordospora",
         Lake=="LilAp",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

lo.spores<- lo$Concentration

res.lo <- wilcox.test(lo6.spores, lo.spores)
res.lo
### not significantly different, W = 394, p = 0.95
```

### LilAp Metschnikowia 6 meters
```{r,message=F,warning=F}

lm6 = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="LilAp",
         Depth =="6")
lm6.spores <- lm6$Concentration

lm = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="LilAp",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

lm.spores<- lm$Concentration

res.lm <- wilcox.test(lm6.spores, lm.spores)
res.lm
### not significantly different, W = 379.5, p = 0.88
```


### LilAp Blastulidium 6 meters
```{r,message=F,warning=F}

lb6 = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="LilAp",
         Depth =="6")
lb6.spores <- lb6$Concentration

lb = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="LilAp",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

lb.spores<- lb$Concentration

res.lb <- wilcox.test(lb6.spores, lb.spores)
res.lb
### not significantly different, W = 349, p = 0.47
```

### LilAp Pasteuria 6 meters
```{r,message=F,warning=F}

lp6 = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="LilAp",
         Depth =="6")
lp6.spores <- lp6$Concentration

lp = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="LilAp",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

lp.spores<- lp$Concentration

res.lp <- wilcox.test(lp6.spores, lp.spores)
res.lp
### not significantly different, W = 471, p = 0.22
```

## LilAp Spirobacillus 6 meters
```{r,message=F,warning=F}

ls6 = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="LilAp",
         Depth =="6")
ls6.spores <- ls6$Concentration

ls = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="LilAp",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5") 

ls.spores<- ls$Concentration

res.ls <- wilcox.test(ls6.spores, ls.spores)
res.ls
### significantly different, W = 637.5, p = 0.0005
```

### Bishop Ordospora top 5 v bottom 1 meter
```{r,message=F,warning=F}

bot = a %>%
  filter(Parasite== "Ordospora",
         Lake=="Bishop",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5")
bot.spores <- bot$Concentration

bob = a %>%
  filter(Parasite== "Ordospora",
         Lake=="Bishop",
         Depth == "16") 

bob.spores<- bob$Concentration

res.bo <- wilcox.test(bot.spores, bob.spores)
res.bo
### not significantly different, W = 434.5, p = .0848
```

### Bishop Blastulidium top 5 v bottom 1 meter
```{r,message=F,warning=F}

bbt = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="Bishop",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5")
bbt.spores <- bbt$Concentration

bbb = a %>%
  filter(Parasite== "Blastulidium",
         Lake=="Bishop",
         Depth == "16") 

bbb.spores<- bbb$Concentration

res.bb <- wilcox.test(bbt.spores, bbb.spores)
res.bb
### not significantly different, W = 404, p = 0.0591
```


### Bishop Metschnikowia top 5 v bottom 1 meter
```{r,message=F,warning=F}

bmt = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="Bishop",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5")
bmt.spores <- bmt$Concentration

bmb = a %>%
  filter(Parasite== "Metschnikowia",
         Lake=="Bishop",
         Depth == "16") 

bmb.spores<- bmb$Concentration

res.bm <- wilcox.test(bmt.spores, bmb.spores)
res.bm
### not significantly different, W = 229, p = 0.0551
```


### Bishop Pasteuria top 5 v bottom 1 meter
```{r,message=F,warning=F}

bpt = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="Bishop",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5")
bpt.spores <- bpt$Concentration

bpb = a %>%
  filter(Parasite== "Pasteuria",
         Lake=="Bishop",
         Depth == "16") 

bpb.spores<- bpb$Concentration

res.bp <- wilcox.test(bpt.spores, bpb.spores)
res.bp
### not significantly different, W = 342.5, p = 0.781
```


### Bishop Spirobacillus top 5 v bottom 1 meter
```{r,message=F,warning=F}

bst = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="Bishop",
         Depth =="0" | Depth == "1" | Depth =="2"| Depth == "3"|Depth == "4"|Depth == "5")
bst.spores <- bst$Concentration

bsb = a %>%
  filter(Parasite== "Spirobacillus",
         Lake=="Bishop",
         Depth == "16") 

bsb.spores<- bsb$Concentration

res.bs <- wilcox.test(bst.spores, bsb.spores)
res.bs
###  significantly different, W = 110, p = 0.0008
```

